---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)

file_url<-"https://esajournals.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1002%2Fecy.2647&file=ecy2647-sup-0001-DataS1.zip"
file_name <- basename(file_url)

if (!dir.exists("bird_data")) dir.create("bird_data")

#if (!file.exists(file_name))
download.file(file_url, destfile = "bird_data/bird_data.zip", mode = "wb")
unzip("bird_data/bird_data.zip", overwrite = TRUE, exdir = "bird_data")

#reading and renaming data
bird_data<-read_csv("bird_data/ATLANTIC_BIRD_TRAITS_completed_2018_11_d05.csv")

#renaming variables 
bird_data <- rename(bird_data, 
                    body_mass_g = Body_mass.g., 
                    body_length_mm = Body_length.mm., 
                    altitude = Altitude, 
                    wing_length_mm = Wing_length.mm.
                    )

#new variable
length_vs_mass <- select(bird_data, body_length_mm, body_mass_g)

length_vs_mass <- mutate(bird_data, 
                         length_mass_ratio = body_length_mm / body_mass_g)


```

# Title

Casey Perkins, Bryn Rivenes

Dr. Chris Merkord

## Introduction

Ecogeographic rules explain variation within a species using biogeography. Allen's rule states that body surface-area-to-volume ratio for animals varies with the average temperature of their habitat. Typically, latitude is used as the temperature variable. However, this rule should hold true over any type of temperature variable. We want to know, is there a correlation between altitude and ratio of mass to length of birds in The Atlantic Forest? To answer this question, we will be using the Atlantic Bird Traits data set. This data set has been compiled over two centuries by tens of researchers (1820-2018) and includes data on 80% of the bird populations found in The Atlantic Forest. We will perform an exploratory analysis on this data set looking for correlations between altitude and ratio of mass to length of bird species. Based on other similar analyses, we expect to see a positive correlation between these variables.

## Methods

where did data come from

show how birds were measured/captured

Using R studio, we created our length to mass ratio data by dividing length by mass for every bird in the Atlantic Bird's Data Set that had a recorded mass and length. We created a histogram for our lmr variable, to test for normal distribution. We found we had a serious right skew. To correct for this, we took the natural log of each data point. Then we separated out each Order into its' own graph and found that Passeriformes made up a disproportionately large part of the data set. We focused in on this order and made a scatterplot of the means of each order, family, genus and species vs altitude. We added a trendline which showed a slight upward trend. We then removed all of the species found at altitudes of less than 1000 and made another scatterplot. This revealed a much sharper positive correlation. Then we performed a t-test comparing means above 1000 and means below 1000 and found (enter what we find)

## Results

one sentence description of conclusion

```{r sestup, include = true}
#ggplot Length Mass Ratio vs Altitude
ggplot(data = length_vs_mass) +
  geom_point(mapping = aes(
    y = length_mass_ratio, 
    x = altitude)
  )

```

```{r}
#testing for normal distribution of length_mass_ratio
ggplot(data = length_vs_mass) +
  geom_histogram(mapping = aes(
    x = ln_lmr), bins = 100)+ 
  facet_wrap(~Order)

```

```{r}
#testing for normal distribution of length_mass_ratio
ggplot(data = length_vs_mass) +
  geom_histogram(mapping = aes(
    x = length_mass_ratio), bins = 100)
#seperate by order
ggplot(data = length_vs_mass) +
  geom_histogram(mapping = aes(
    x = length_mass_ratio), bins = 100) +
  facet_wrap(~Order)
```

```{r}
#log of length_mass_ratio to correct right skew 
length_vs_mass <-mutate(length_vs_mass, ln_lmr = log(length_mass_ratio))

#testing for normal distribution of ln_lmr
ggplot(data = length_vs_mass) +
  geom_histogram(mapping = aes(
    x = ln_lmr), bins = 100)+ 
  facet_wrap(~Order)
```

```{r}
length_vs_mass %>% 
  filter(
    Order=="Passeriformes",
    Family=="Furnariidae",
  ) %>% 
  ggplot(mapping = aes(
    y = ln_lmr, 
    x = altitude)) +
  geom_point() +
  geom_smooth(method = "lm")

length_vs_mass %>% 
  filter(
    Order=="Passeriformes",
    Family=="Furnariidae",
    altitude > 1000
  ) %>% 
  ggplot(mapping = aes(
    y = ln_lmr, 
    x = altitude)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Discussion

We can see that there is a correlation between the mean lmr of bird species within order Passeriformes and the altitude at which they are found. Our graph suggests that the greater the altitude a bird species is found at, the greater their mean lmr is. This trend follows the trend seen in birds examined latitudinally as seen in many other research projects on birds. Essentially, what we see is that birds that are fat and fluffy tend to live at higher elevations that birds that are long and skinny. However, it is important to note that this trend was far more apparent in birds over 1000 in elevation. There are a couple of possible explanations for this. Generally, at higher elevations, there are colder temperatures, so birds require a higher body to length ratio to stay warm. However, birds from birds found in more extreme climates, we would expect to see greater variation in morphology. For example, many bird species that live in elevations below 1000 can be found at 100, 500, or 800 easily. However, birds that are found in more extreme climates (higher elevations) tend to be more adapted for their specific environment. Allen's rule suggest that any morphological trait (such as appendage thickness or hair length) should also vary based on temperature. A future analysis we would like to do is see if this trend holds true for other morphological bird characteristics.  

## References

-   Blackburn T, Ruggiero A. Latitude, elevation and body mass variation in Andean passerine birds. Global Ecology and Biography 2001 (cited 2022 Feb 18); 10: 245-259.

-   Hasui, É., Metzger, J.P., Pimentel, R.G., Silveira, L.F., Bovo, A.A.d.A., Martensen, A.C., Uezu, A., Regolin, A.L., Bispo de Oliveira, A.Â., Gatto, C.A.F.R., Duca, C., Andretti, C.B., Banks-Leite, C., Luz, D., Mariz, D., Alexandrino, E.R., de Barros, F.M., Martello, F., Pereira, I.M.d.S., da Silva, J.N., Ferraz, K.M.P.M.d.B., Naka, L.N., dos Anjos, L., Efe, M.A., Pizo, M.A., Pichorim, M., Gonçalves, M.S.S., Cordeiro, P.H.C., Dias, R.A., Muylaert, R.d.L., Rodrigues, R.C., da Costa, T.V.V., Cavarzere, V., Tonetti, V.R., Silva, W.R., Jenkins, C.N., Galetti, M. and Ribeiro, M.C. (2018), ATLANTIC BIRDS: a data set of bird species from the Brazilian Atlantic Forest. Ecology, 99: 497-497. <https://doi.org/10.1002/ecy.2119>

-   Lamichhaney S, Han F, Berglund J, Wang C, Almén MS, Webster MT, Grant BR, Grant PR, Andersson L. A beak size locus in Darwin's finches facilitated character displacement during a drought. Science. 2016 (cited 2022 Feb 18);352(6284):470-4

-   Landmann A, Winding N. Guild Organization and Morphology of High-Altitude Granivorous and Insectivorous Birds: Convergent Evolution in an Extreme Environment. Oikos. 2995 (cited 2022 Feb 18);73(2):237-250.

-   Sun Y, Li M, Song G, Lei F, Li D, Wu Y. The Role of Climate Factors in Geographic Variation in Body Mass and Wing Length in a Passerine Bird. Avian Res. 2017 (cited 2022 Feb 19); 8, 1.

## Acknowledgments

We would like to thank the original authors of the Atlantic Birds Data Set and all of those who helped compile it through museum collections, online data bases, published sources, and unpublished reports, as well as the people out out in the field who originally collected this data.

We also would like to thank Dr. Merkord for his guidance with coding and data analyse.
